---
title: "linear models"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)

set.seed(1)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

**linear regression lm()** 
output must be continuous but the predictor can be any type, if the predictor is categorical, it require indicator variable(1/0)

betw1: expected change in y for one unit change in x, keeping sex fixed

test statistic:
t=(beta_hat-beta)/se_hat(beta_hat)
test multiple coefficients by F test or ANOVA
**Generalized linear models: logictic regression**
for non-continuius outcomes  



## Import data

```{r}
data("nyc_airbnb")
nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(stars = review_scores_location / 2) %>% 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) %>% 
  filter(borough != "Staten Island") %>% 
  select(price, stars, borough, neighborhood, room_type)
```

## Fit a model
how the price is related to stars and borough
```{r}
nyc_airbnb %>% 
  ggplot(aes(x = stars, y = price, color = borough)) +
  geom_point()
```

Let's fit the model we care about.

```{r}
#brones are the reference since it in  alphabetical order, 
#  specify an intercept-only model (outcome ~ 1), a model with no intercept (outcome ~ 0 + ...), and a model using all available predictors (outcome ~ .).
fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

Let's look at the result...

```{r, include = FALSE}
# result of lm() is list
fit
summary(fit)
summary(fit)$coef 
coef(fit)
residuals(fit)
```

Let's look at the result better ..

```{r}
# both the glance and tudy produce data fram but not list
broom::glance(fit)# a glance of the result
broom::tidy(fit) %>% 
  select(-std.error, -statistic) %>% 
  mutate(
    term = str_replace(term, "borough", "Borough: ")
  ) %>% 
  knitr::kable(digits = 3)
```

## Be in control of factors

```{r}
#the factor level is treated as the reference. 
#it’s important to note that changing reference categories won’t change “fit” or statistical sigificance, but can affect ease of interpretation
nyc_airbnb =
  nyc_airbnb %>% 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type)
  )
```

Look at that plot again


```{r}
# now borough is in an order of frequency 
# now manhattan is the reference
nyc_airbnb %>% 
  ggplot(aes(x = stars, y = price, color = borough)) + 
  geom_point()
```

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
broom::tidy(fit)
broom::glance(fit)
```

## Diagnostics

```{r}
#add residuals and fitted values to a dataframe.
nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = borough, y = resid)) + 
  geom_violin() +
  ylim(-500, 1500)
# ylim() can help to zoom in, delet the big outlier :notably the presence of extremely large outliers in price and a generally skewed residual distribution. 


nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = stars, y = resid)) +
  geom_point() + 
  facet_wrap(. ~ borough)
# manhattan get wider distribution of residule whe nyou get higher star
# Bronx has small range of residuals in different starts
```

## Hypothesis tests

This does t-test by default

```{r}
fit %>% 
  broom::tidy()
```

what about the significance of `borough`.

```{r}
# in the null model the price is unassociated with  stars
# the hypothesis model th price is related to borough
fit_null = lm(price ~ stars, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough, data = nyc_airbnb)
anova(fit_null, fit_alt) %>% 
  broom::tidy()
```


## Nest data, fit models

This is pretty formal and also complex

```{r}
# how star affect the price when this borough compared to that borough
#how room_type affect the price when this borough compared to that borough
# in the term column: the 'star' means the the overall effect of stars
# the combination of stars and borough, stars:boroughBrooklyn mean the effect of stars in a different goup
# what is going on in my reference group of mahaten if I want to know waht effect room type 
fit = lm(price ~ stars * borough + room_type * borough, data = nyc_airbnb)
broom::tidy(fit)
```

This is more exploratory but maybe easier to understand.

```{r}
# map linear model across data(it is a data frame)
# you will get one column named model that contains 4 linear model fits
# map broom:tidy across model column and get a new column named results that contains 4 lists of data frame
# the lists of data frame can be unnest 
nyc_airbnb %>% 
  nest(data = -borough) %>% # nest everything except borough
  mutate(
    models = map(.x = data, ~lm(price ~ stars + room_type , data = .x)), 
    results = map(models, broom::tidy)
  ) %>% 
  select(-data, -models) %>% 
  unnest(results) %>% 
  filter(term != "(Intercept)") %>% 
  select(borough, term, estimate) %>% 
  pivot_wider(
    names_from = borough, 
    values_from = estimate
  )
```



Let's nest even more ....
  
  shows neighborhood-specific estimates for the coefficients related to room type.
```{r}
# map(): for each of the models, fit a model that relates price to stars and room_type using the data frame
# the estimate is negative: as stars increase, the price in  chinatown will decrease
nyc_airbnb %>% 
  filter(borough == "Manhattan") %>% 
  nest(data = -neighborhood) %>% 
  mutate(
    models = map(.x = data, ~lm(price ~ stars + room_type , data = .x)), 
    results = map(models, broom::tidy)
  ) %>% 
  select(-data, -models) %>% 
  unnest(results) %>% 
  filter(str_detect(term, "room_type")) %>% 
  ggplot(aes(x = neighborhood, y = estimate)) +
  geom_point() +
  facet_wrap(. ~ term) + 
  theme(axis.text.x = element_text(angle = 70, vjust = 1, hjust = 1))

#map everything against the term
```

the plot above shows that a reduction in room price for a private room or a shared room compared to an entire apartment, but this varies quite a bit across neighborhoods.

  
  
 it really isn’t a good idea to fit models with main effects or interactions for each. Instead, you’d be best-off using a mixed model, with random intercepts and slopes for each neighborhood. 

manhattan_airbnb =
  nyc_airbnb %>% 
  filter(borough == "Manhattan") %>% 
  lme4::lmer(price ~ stars + room_type + (1 + room_type | neighborhood), data = .) %>% 
  broom.mixed::tidy()


